---
title: "NY Squirrels"
author: "Caroline Carrozzini"
date: "1/31/2023"
output: html_document
---

```{r}
install.packages("tidytuesdayR")
library(tidyverse)
```
```{r}
nyc_squirrels <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-29/nyc_squirrels.csv")
```

```{r}
install.packages("tidymodels")
library(tidymodels)
nyc_squirrels_splits <- initial_split(nyc_squirrels, prop = 0.5)
exploratory_data <- training(nyc_squirrels_splits)
test_data <- testing(nyc_squirrels_splits)
```


# Exploratory Data Questions

+ What was the distribution of squirrel color?

+ Where are the squirrels found?

+ How many of the squirrels that were kukking were also tail twitching or tail flagging?

+ How many of the squirrels that were running were also running away from something?

```{r}
nyc_squirrels %>%
  ggplot() + 
  geom_bar(mapping = aes(x = primary_fur_color)) + 
  labs(title = "Counts of Squirrel Fur Color Distribution", 
       x = "Fur Color", y = "Count")
```

```{r}
nyc_squirrels %>%
  ggplot() +
  geom_point(mapping = aes(x = long, y = lat)) +
  labs(title = "Location of Squirrels",
       x = "Longitude",
       y = "Latitude")
```

```{r}
nyc_squirrels %>%
  ggplot() +
  geom_bar(mapping = aes(x = age, fill = primary_fur_color)) +
  labs(title = "Fur color by age",
       x = "Age",
       y = "Fur color")
```

```{r}
nyc_squirrels %>%
  ggplot() +
  geom_point(mapping = aes(x = long, y = lat, color = primary_fur_color)) +
  labs(title = "Location of Squirrels",
       x = "Longitude",
       y = "Latitude")
```

```{r}
nyc_squirrels %>%
  ggplot() +
  geom_density(mapping = aes(x = long),
                 color = "white",
                 fill = "blue") +
  labs(title ="Distribution of Longitude",
       x = "Longitude)", y = "Count")
```

```{r}
nyc_squirrels %>%
  ggplot() +
  geom_density(mapping = aes(x = lat),
                 color = "white",
                 fill = "blue") +
  labs(title ="Distribution of Latitude",
       x = "Latitude", y = "Count")
```

```{r}
nyc_squirrels %>%
  ggplot() +
  geom_point(mapping = aes(x = long, y = lat, color = eating)) +
  labs(title = "Location of Squirrels",
       x = "Longitude",
       y = "Latitude")
```

```{r}
nyc_squirrels %>%
  filter(above_ground_sighter_measurement != FALSE) %>%
  mutate(above_ground_sighter_measurement= as.numeric(above_ground_sighter_measurement)) %>%
  mutate(age= ifelse(age=="?", NA, age)) %>%
  ggplot() +
  geom_density(mapping = aes(x = above_ground_sighter_measurement, fill= age),
                 color = "white",
                 fill = "blue") +
  labs(title ="Above Ground Sightings",
       x = "Height", y = "Count") +
  facet_wrap(~age, ncol = 1)
```

# Hypotheses

+ 1. Do squirrels fur colors change with age?
+ 2. Do squirrels climb higher up with an increase in age?